---
title: "STRs in San Antonio’s Eastside"
author: "Esteban Lopez Ochoa & Ashley Leal"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background

Sort Term Rentals (STRs) are increasingly becoming an important source of accomodation for tourism. As such, cities have experienced the benefits from the expansion of STRs in their tourism-related industries, however, there is also a potential side effect from this activity that comes in the form of a higher housing demand, less ownership and renting housing stock for long term residents, and ultimately gentrification. This article explores the connection between STRs evolution and gentrification/displacement pressures using two neighborhoods in San Antonio's Eastside.


## STR data

```{r data,message=FALSE,echo=FALSE,warning=FALSE}
library(data.table)
library(readxl)
srt_202<-data.table(read_xlsx("str_data/78202 STRs.xlsx"),stringsAsFactors = F)
srt_203<-data.table(read_xlsx("str_data/78203 STRs.xlsx"),stringsAsFactors = F)

STR<-rbind(srt_202,srt_203)
```


This analysis focuses on two zipcodes `78202` and `78203`. Figure 1 bellow shows the spatial distribution of STR permits currently active in the City of San Antonio.  

```{r Fig1, echo=FALSE}
library(sf)
STR<-STR[!is.na(X),]
STR<-st_as_sf(x = STR,coords=c("X","Y"),crs=3674)
library(mapview)
mapview(STR,zcol="ZipCode")
```

# STR spatial density

```{r Sp2D,echo=FALSE,message=FALSE}
library(ggplot2)

library(tigris)
bexar_tracts<-tracts(state = "TX",county = "Bexar",cb=T)
eastside_tracts<-bexar_tracts[bexar_tracts$TRACTCE%in%c("191900","110300","130200","130600","130500","140100","130300","130401","130402"),]

ggplot()+
  geom_sf(data=eastside_tracts)+
  geom_sf(data = STR)+
  stat_density_2d(data = STR, 
                  mapping = aes(x = purrr::map_dbl(geometry, ~.[1]),
                                y = purrr::map_dbl(geometry, ~.[2]),
                                fill = after_stat(density)),
                  geom = 'tile',
                  contour = FALSE,
                  alpha = 0.8)+
  scale_fill_viridis_c(option = 'magma', direction = -1)
```

